Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-22 paper/v2/paper-edited.tex
edits

\textbf{Abstract---}
We develop \sysname, a language and compiler to help network operators with a challenging, error-prone task---bridging the gap between network-wide routing objectives and low-level configurations of devices that run complex, distributed protocols.
%
The language allows operators to specify their objectives naturally, using 
high-level constraints on both the shape and relative preferences of traffic paths.
%
The compiler automatically translates these specifications to router-level BGP configurations, using an effective intermediate representation that compactly encodes the flow of routing information along policy-compliant paths.
%It uses We introduce data structures and algorithms that
It guarantees that the compiled configurations correctly implement the specified policy under all possible combinations of failures.
%
We show that \sysname can effectively express the policies of datacenter and backbone networks of a large cloud provider;  and despite its strong guarantees, our compiler scales to networks with hundreds or thousands of routers.

@@

We have implemented our language and compiler in a system called \sysname. To evaluate it, we use it to specify real policies of datacenter and backbone networks.
% and compared these specifications with existing configurations.
We find that our language expresses such policies easily, and that the compiler scales to topologies with hundreds to thousands of routers, compiling in under 9 minutes in all cases.

@@

To implement policy P1, the operators must compute and assign
local preferences such that preferences at \CD{Cust}-facing interfaces
$>$ \CD{Peer}-facing interfaces $>$ \CD{Prov}-facing interfaces. At
the same time, to satisfy P3, the preference at \CD{R2}'s
\CD{Cust}-facing interface should be lower than that at
\CD{R1}. Implementing P3 will also require MEDs to be appropriately configured on \CD{R1} and \CD{R2}.
To implement P2, the operators can assign communities that
indicate where a certain routing announcement entered the
network. Then, \CD{R4} must be configured to not announce to
$\CD{Peer}$ routes that have communities that correspond to the
\CD{R2}-\CD{Prov} link but to announce routes with communities for the \CD{R2}-\CD{Cust} and \CD{R1}-\CD{Cust} links. A similar policy must be configured for \CD{R2} as well. Finally, to implement P4 and P5, the operators will have to compute and configure appropriate prefix- and AS-path-based import and export filters at each router.

@@

\sysname lets operators configure the network with the
abstraction that they have centralized control over routing.
Specifically, the operator simply provides a set of high-level constraints
that describe the
paths traffic should---or should not---take and their relative preferences.
\sysname specifications are written modularly via a series
of declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network from the previous section, we first express the idea that
we prefer that traffic leave the network through \CD{R1} over \CD{R2} (to \CD{Cust}) over \CD{Peer} over \CD{Prov} (policy P1 and P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define Prefs = \Exit(R1 \Prefer R2 \Prefer Peer \Prefer Prov)
\end{code}
This  statement defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) for which traffic exits our network
through either router \CD{R1}, router \CD{R2}, \CD{Peer}, or \CD{Prov}.
The paths that exit through \CD{R1}
are preferred (\Prefer) to those that exit through \CD{R2}, which are preferred to those that
leave through \CD{Peer} and then \CD{Prov}.  As we
describe in the next section, the \Exit\ predicate, as well as other
path predicates used later in this section, is simply a shorthand for
a particular regular expression over paths that is expressible in our policy
language. The preference operator (\Prefer) is flexible and can be used between constraints as well as among individual routers. For example, the above constraint could have been written equivalently as
\Exit(\CD{R1}) \Prefer \ldots \Prefer ~\Exit(\CD{Prov})

@@

     pol     &::=& p_1, \dots, p_n & \textit{constraints} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m & \textit{preferences} \\
     x       &::=& d.d.d.d/d & \textit{prefix} \\
     t       &::=& \True & \textit{true} \\
         &\BNFALT& \NOT t & \textit{negation} \\
         &\BNFALT& t_1 \OR{} t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \AND{} t_2 & \textit{conjunction} \\
         &\BNFALT& \textit{prefix} = x & \textit{prefix test} \\
         &\BNFALT& \textit{comm} = d & \textit{community test} \\
     r       &::=& l & \textit{location} \\
         &\BNFALT& \emptyset & \textit{empty set} \\
         &\BNFALT& \In & \textit{internal loc} \\
         &\BNFALT& \Out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& \NOT r & \textit{path negation} \\
         &\BNFALT& r^* & \textit{iteration} \\
     lp       &::=& r_1 \rightarrow r_2 & \textit{link pairs} \\
     cc     &::=& agg(x, lp) \BNFALT tag(c, t, lp) & \textit{control constraints} \\
  \end{array} \]

@@

To handle these challenges, we decompose compilation into multiple stages, shown in  Figure~\ref{fig:pipeline}, and develop efficient algorithms for the translation between stages. The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked for being well-formed (\EG, never constraining traffic that does not enter the network), before being combined with the network topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines safety, it translates the PGIR to an abstract BGP (ABGP) representation. ABGP can then be translated into various vendor-specific device
configurations as needed.
%The \sysname compiler currently generates
% Quagga router configurations from ABGP.

@@

\sysname also supports constraints purely on the control-plane behavior of BGP. For example, prefix aggregation is an important optimization to reduce routing table size and churn in practice. A constraint of the form $agg(x,l)$ tells the compiler to perform aggregation for prefix $x$ across all links described by $l$. It is also often useful to be able to add community tags to exported routes in BGP (\EG, to communicate non-standard information to peers). A constraint of the form $tag(c,t,l)$ adds community tag $c$ for any prefixes matching $t$ across links $l$.
%% Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
%% $$\Agg(128.17.0.0/16, \In \Link \Out)$$
%% where the expression $\In \Link Out$ refers to control messages flowing from any internal to any external location.
We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but we also support other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.

@@

Now that the user policy exists in a simplified form, we must consider the topology. In particular, we want a compact representation that describes all the possible ways BGP route announcements can flow through the network subject to the policy and topology constraints. The PGIR captures these constraints by ``intersecting" each of the regular automata corresponding to the RIR path preferences and the topology. Paths through the PGIR correspond to real paths through the topology that satisfy the user constraints.

@@

% Formally define automata
%The RIR policy is an ordered sequence of regular expressions $r_1 \Prefer \dots \Prefer r_j$.
While paths in an RIR policy describe the direction traffic flows through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated, \IE, route announcements flowing in the opposite direction. To capture this idea, for each regular expression $r_i$ in RIR policy, we construct a deterministic finite state machine on the reversed regular expression. Each automaton for the reversed regular expression of $r_i$ is a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all abstract locations (\IE, routers or ASes), $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$ and a set of directed edges $E \colon V \times V$.
%
% Formally define product graph
The combined PGIR is a tuple ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of preference ranks.

@@

Throughout the remainder of the paper, we will the convention that metavariables $m$ and $n$ stand for PGIR nodes and $l$ stands for a topology location. Captial letters like $X$, $Y$ refer to concrete topology locations, while capital letters with subscripts such as $X_1$ and $X_2$ refer to concrete PGIR nodes that share a topology location (\IE, $\tilde{X_1}$ = $\tilde{X_2} = X$).

@@

Intuitively, a PGIR tracks the policy states of each automaton as route announcements move between locations.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from neighbor W that allows traffic to enter the network at $A$ and utilize the C--D link before leaving the network (through $X$ or $Y$). As a backup, we also want to allow traffic to enter the network from $B$, in which case the traffic can also utilize the C--E link before leaving the network. For simplicity, we assume that the route ends in either $X$, $Y$, or $Z$. The RIR for the policy could be written as:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in the PGIR corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $n$ such that the preference function $P(n) = \set{i_1, \dots, i_k}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_k$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining a path with the lowest (best) rank of $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This possibility is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton cannot accept the current path or any extension of it. Since
node $(W,5,-)$ is in an accepting state for the first automaton, it indicates that this path has preference 1.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=.66\columnwidth]{figures/loops}
  \caption{A Product Graph where preference inference is unsound before considering loops. Path ranks shown by nodes.}
  \label{fig:avoiding-loops}
  %\vspace{-1em}
\end{figure}

>>>>

\textbf{Abstract---}
We develop \sysname, a language and compiler to help network operators with a challenging, error-prone task---bridging the gap between network-wide routing objectives and low-level configurations of devices that run complex, distributed protocols.
%
The language allows operators to specify their objectives naturally, using 
high-level constraints on both the shape and relative preference of traffic paths.
%
The compiler automatically translates these specifications to router-level BGP configurations, using an effective intermediate representation that compactly encodes the flow of routing information along policy-compliant paths.
%It uses We introduce data structures and algorithms that
It guarantees that the compiled configurations correctly implement the specified policy under all possible combinations of failures.
%
We show that \sysname can effectively express the policies of datacenter and backbone networks of a large cloud provider;  and despite its strong guarantees, our compiler scales to networks with hundreds or thousands of routers.

@@

We have implemented our language and compiler in a system called \sysname. To evaluate it, we use it to specify real policies for Microsoft's datacenter and backbone networks.
% and compared these specifications with existing configurations.
We find that our language expresses such policies easily, and that the compiler scales to topologies with hundreds or thousands of routers, compiling in under 9 minutes in all cases.

@@

To implement policy P1, the operators must compute and assign
local preferences such that preferences at \CD{Cust}-facing interfaces
$>$ \CD{Peer}-facing interfaces $>$ \CD{Prov}-facing interfaces. At
the same time, to satisfy P3, the preference at \CD{R2}'s
\CD{Cust}-facing interface should be lower than that at
\CD{R1}. Implementing P3 will also require MEDs to be appropriately configured on \CD{R1} and \CD{R2}.
To implement P2, the operators can assign communities that
indicate where a certain routing announcement entered the
network. Then, \CD{R4} must be configured to not announce to
$\CD{Peer}$ routes that have communities that correspond to the
\CD{R2}-\CD{Prov} link but to announce routes with communities for the \CD{R2}-\CD{Cust} and \CD{R1}-\CD{Cust} links. A similar type of policy must be configured for \CD{R2} as well. Finally, to implement P4 and P5, the operators will have to compute and configure appropriate prefix- and AS-path-based import and export filters at each router.

@@

\sysname lets operators configure the network with the
abstraction that they have centralized control over routing.
Specifically, the operator simply provides a set of high-level constraints
that describe the
paths traffic should---or should not---take and their relative preferences.
\sysname specifications are written modularly via a series
of declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network from the previous section, we first express the idea that
we prefer that traffic leave the network through \CD{R1} over \CD{R2} (to \CD{Cust}) over \CD{Peer} over \CD{Prov} (policy P1 and P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define Prefs = \Exit(R1 \Prefer R2 \Prefer Peer \Prefer Prov)
\end{code}
This  statement defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) for which traffic exits our network
through either router \CD{R1}, router \CD{R2}, \CD{Peer}, or \CD{Prov}.
The paths that exit through \CD{R1}
are preferred (\Prefer) to those that exit through \CD{R2}, which are preferred to those that
leave through \CD{Peer} and then \CD{Prov}.  As we
describe in the next section, the \Exit\ expression, as well as other
path expressions used later in this section, is simply a shorthand for
a particular regular expression over paths that is expressible in our policy
language. The preference operator (\Prefer) is flexible and can be used between constraints as well as among individual routers. For example, the above constraint could have been written equivalently as
\Exit(\CD{R1}) \Prefer \ldots \Prefer ~\Exit(\CD{Prov})

@@

     pol     &::=& p_1, \dots, p_n & \textit{constraints} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m & \textit{preferences} \\
     x       &::=& d.d.d.d/d & \textit{prefix} \\
     t       &::=& \True & \textit{true} \\
         &\BNFALT& \NOT t & \textit{negation} \\
         &\BNFALT& t_1 \OR{} t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \AND{} t_2 & \textit{conjunction} \\
         &\BNFALT& \textit{prefix} = x & \textit{prefix test} \\
         &\BNFALT& \textit{comm} = d & \textit{community test} \\
     r       &::=& l & \textit{location} \\
         &\BNFALT& \emptyset & \textit{empty set} \\
         &\BNFALT& \In & \textit{internal loc} \\
         &\BNFALT& \Out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& \NOT r & \textit{path negation} \\
         &\BNFALT& r^* & \textit{iteration} \\
     lp       &::=& r_1 \rightarrow r_2 & \textit{link pairs} \\
     cc     &::=& agg(x, lp) \BNFALT tag(d, t, lp) & \textit{control constraints} \\
  \end{array} \]

@@

To handle these challenges, we decompose compilation into multiple stages, shown in  Figure~\ref{fig:pipeline}, and develop efficient algorithms for the translation between stages. The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked well-formedness (\EG, never constraining traffic that does not enter the network), before being combined with the topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines safety, it translates the PGIR to an abstract BGP (ABGP) representation. ABGP can then be translated into various vendor-specific device
configurations as needed.
%The \sysname compiler currently generates
% Quagga router configurations from ABGP.

@@

\sysname also supports constraints purely on the control-plane behavior of BGP. For example, prefix aggregation is an important optimization to reduce routing table size in practice. A constraint of the form $agg(x,lp)$ tells the compiler to perform aggregation for prefix $x$ across all links described by $lp$. It is also often useful to be able to add community tags to exported routes in BGP (\EG, to communicate non-standard information to peers). A constraint of the form $tag(d,t,lp)$ adds community tag $d$ for any prefixes matching $t$ across links $lp$.
%% Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
%% $$\Agg(128.17.0.0/16, \In \Link \Out)$$
%% where the expression $\In \Link Out$ refers to control messages flowing from any internal to any external location.
We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but \sysname also supports other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.

@@

Now that the user policy exists in a simplified form, we must consider the topology. In particular, we want a compact representation that describes all the possible ways BGP route announcements can flow through the network subject to the policy and topology constraints. The PGIR captures these constraints by ``intersecting" each of the regular automata corresponding to the RIR path preferences with the topology. Paths through the PGIR correspond to real paths through the topology that satisfy the user constraints.

@@

% Formally define automata
%The RIR policy is an ordered sequence of regular expressions $r_1 \Prefer \dots \Prefer r_j$.
While paths in an RIR policy describe the direction traffic flows through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated, \IE, route announcements flowing in the opposite direction. To capture this idea, for each regular expression $r_i$ in an RIR policy, we construct a deterministic finite state machine on the reversed regular expression. Each automaton for the reversed regular expression of $r_i$ is a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all abstract locations (\IE, routers or ASes), $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$ and a set of directed edges $E \colon V \times V$.
%
% Formally define product graph
The combined PGIR is a tuple ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of path ranks.

@@

Throughout the remainder of the paper, we will use the convention that metavariables $m$ and $n$ stand for PGIR nodes and $l$ stands for a topology location. Capital letters like $X$ refer to concrete topology locations, while capital letters with subscripts such as $X_1$ and $X_2$ refer to concrete PGIR nodes that share a topology location (\IE, $\tilde{X_1}$ = $\tilde{X_2} = X$).

@@

Intuitively, the PGIR tracks the policy states of each automaton as route announcements move between locations.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from neighbor W that allows traffic to enter the network at $A$ and utilize the C--D link before leaving the network (through $X$ or $Y$). As a backup, we also want to allow traffic to enter the network from $B$, in which case the traffic can also utilize the C--E link before leaving the network. For simplicity, we assume that the route ends in either $X$, $Y$, or $Z$. The RIR for the policy could be written as:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in the PGIR corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $n$ such that the preference function $P(n) = \set{i_1, \dots, i_k}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_k$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining a path with the lowest (best) rank of $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This possibility is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton cannot accept the current path or any extension of it. Since
node $(W,5,-)$ is in an accepting state for the first automaton, it indicates that this path has preference 1.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=.67\columnwidth]{figures/loops}
  \caption{A Product Graph where preference inference is unsound before considering loops. Path ranks shown by nodes.}
  \label{fig:avoiding-loops}
  %\vspace{-1em}
\end{figure}
